---
title: "P8105_hw6_yw4200"
author: "yh wang"
date: "2023-11-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.

Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
library(dplyr)
homicide = read.csv("./local data/homicide-data.csv")

# create a city_state variable
homicide_tidy <- homicide |>
  mutate(city_state = paste(city, state, sep = ", "),
         solved_binary = ifelse(disposition == "Closed by arrest",1,0),
         victim_age = as.numeric(victim_age)) |>
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) |>
  filter(victim_race == "White" | victim_race == "Black")
  
summary(homicide_tidy)
```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.
```{r}
# prepare the Baltimore, MD data for glm
glm_data <- homicide_tidy |>
  filter(city == "Baltimore") |>
  select(solved_binary, victim_age, victim_race, victim_sex) %>%
  mutate(victim_sex = as.factor(victim_sex))

# glm regression
glm_model1 <- glm(solved_binary ~ victim_age + victim_race + victim_sex, data = glm_data, family = binomial())

save(glm_model1, file = "./result/baltimore_logistic_regression_model.RData")

glm_model1 |>
  broom::tidy() |> 
  knitr::kable(digits = 3)

# obtain the estimate and confidence interval of the adjusted odds ratio
glm_model1 |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate),
         OR_lower = exp(estimate - 1.96 * std.error),
         OR_upper = exp(estimate + 1.96 * std.error)) |>
  select(term, log_OR = estimate, OR, OR_lower, OR_upper) |> 
  filter(term == 'victim_sexMale') |>
  knitr::kable(digits = 3)

```

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.
```{r}
library(tidyr) 
library(purrr)
# fit logistic regression for all city
city_model <- homicide_tidy |>
   nest(data = -city_state) |>
  mutate(
    fit = map(.x = data, ~ glm(solved_binary ~ victim_age + victim_sex + victim_race, data = ., family = binomial())),output = map(fit, broom::tidy)
  ) |>
  unnest(cols = output) |>
  mutate(citystate = city_state,
         OR = exp(estimate),
         OR_lower = exp(estimate - 1.96 * std.error),
         OR_upper = exp(estimate + 1.96 * std.error)) |>
  select(citystate, term, log_OR = estimate, OR, OR_lower, OR_upper) |> 
  filter(term == 'victim_sexMale') |>
  arrange(desc(OR))

city_model |> knitr::kable(digits = 3)
```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
```{r}
library(forcats)
library(ggplot2)
ggplot(city_model, aes(x = fct_reorder(citystate, OR), y= OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = OR_lower, ymax = OR_upper)) +
  coord_flip() +
  labs(x = 'citystate', y = 'adjusted odds ratio', title = 'Estimated ORs and CIs by city') +
  theme(axis.text.y = element_text(hjust = 1,size = 5))
```

## Problem 2
For this problem, we’ll use the Central Park weather data similar to data we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

head(weather_df)
```
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1). Note: broom::glance() is helpful for extracting r̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 1∗β̂ 2)

```{r}
# create 5000 bootstrap samples and produce estimates
set.seed(123)

bootstrap_estimates <- weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(model = map(strap, ~lm(tmax ~ tmin +prcp, data = .x)),
         result_tidy= map(model, broom::tidy),
         result_glance = map(model, broom::glance)) |>
  unnest(result_tidy,result_glance) |>
  select(.id, term, estimate, r.squared) |>
  pivot_wider(names_from = term, values_from = estimate) |>
  rename(
    beta_0 = `(Intercept)`,
    beta_1 = tmin,
    beta_2 = prcp) 

result_boot = bootstrap_estimates |>
  mutate(r.squared = r.squared,
    log_beta12 = log(beta_1 * beta_2))

result_filter = result_boot |>
  na.omit()

head(result_boot)
head(result_filter)
  
```
```{r}
# identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval

# CI for R_2 
  quantile(pull(result_boot,r.squared), probs = c(0.025,0.975),na.rm = TRUE) |>
  knitr::kable(digits = 5, col.names = "R_squared")

# CI for R_2 without NA
  quantile(pull(result_filter,r.squared), probs = c(0.025,0.975),na.rm = TRUE) |>
  knitr::kable(digits = 5, col.names = "R_squared")
# CI for log(beta1*beta2)
  quantile(pull(result_filter,log_beta12), probs = c(0.025,0.975),na.rm = TRUE) |>
  knitr::kable(digits = 5,  col.names = "log(beta1*beta2)")
```
```{r}
# Plot the distribution of estimates
result_boot |>
  ggplot(aes(x = r.squared)) + 
  geom_density() +
  labs(
    title = "Distribution of Estimates of R_squared",
    x = "Estimates of R_squared",
    y = "Density"
  )
```
Comment: The distribution of R_squared is approximately normal distribution but a little left-skewed. The value of R_squared is close to 0.92, which means there are strong linear relation between tmax and its predictors(tmin and prcp). It indicates that approximately 92% of the variability in the response variable `tmax` can be explained by predictor variables `tmin` and `prcp` included in the regression model.

```{r}
# Plot the distribution of log(beta1*beta2)
result_filter |>
  ggplot(aes(x = log_beta12)) + 
  geom_density() +
  labs(
    title = "Distribution of Estimates of log(beta1*beta2)",
    x = "Estimates of log(beta1*beta2)",
    y = "Density"
  )
```
Comment:
The distribution is left-skewed with the mean of `r mean(result_filter$log_beta12)` and the sd of `r sd(result_filter$log_beta12)`. There are some outliers, so the large samples are useful.

## Problem 3
In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables.

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```{r}
# import and clean data
birthweight = read.csv("./local data/birthweight.csv")

# check for missing data
missing_data <- sapply(birthweight, function(x) sum(is.na(x)))
missing_data

# convert some numeric into factor and label them
birth_clean <- birthweight |> 
  janitor::clean_names() |>
  mutate(babysex = factor(babysex, levels = c(2, 1), 
                     labels = c("Female", "Male")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), 
                   labels = c("White", "Black", "Asian", 
                              "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present")))

head(birth_clean)
```
Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}
# Propose a regression model for birthweight
# First consider a model with the main effects of all variables and then use stepwise regression to select appropriate variables

# full model
model_all <- lm(bwt ~ ., data = birth_clean)

# stepwise regression
model_stepwise <- step(model_all, direction = "backward")

model_all |> 
  broom::tidy() |>
  knitr::kable(digits = 3)

summary(model_stepwise)

model_stepwise |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)
```

```{r}
# The Variance Inflation Factor (VIF) is a common indicator used to detect multicollinearity in regression models. A higher VIF value indicates that the corresponding independent variable is more likely to be influenced by other independent variables. Typically, VIF values exceeding 10 or 20 are considered indicative of the presence of multicollinearity.
library(car)
vif_values <- vif(model_stepwise)
print(vif_values)
# The result is no multicollinearity.
```

Show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}
library(modelr)
birth_clean %>% 
    add_predictions(model_stepwise) %>% 
    add_residuals(model_stepwise) %>% 
    ggplot(aes(x = pred, y = resid)) +
    geom_point()  + 
    geom_smooth(method = "lm") + 
    labs(title = "Model residuals against fitted values", 
       x = "Predicted values", 
       y = "Residuals") +
  theme_minimal()
```
Comment:

We can see that the slope of the line is close to zero, which is difficult for us to find the relationship between the response variable and predictors.

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
option1 <- lm(bwt ~ blength + gaweeks, data = birth_clean)

option1 |>
  broom::tidy() |>
  knitr::kable(digits = 3)

option2 <- lm(bwt ~ bhead * blength * babysex, data = birth_clean)

option2 |>
 broom::tidy() |>
 knitr::kable(digits = 3)
```

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.
```{r}
cv_df = 
  crossv_mc(birth_clean, 100) 

cv_df <- cv_df |>
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble)) |>
 mutate(
    stepwise_mod  = map(train, ~model_stepwise),
    option1_mod  = map(train, ~option1),
    option2_mod  = map(train, ~option2)) |> 
  mutate(
    rmse_stepwise = map2_dbl(stepwise_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_option1 = map2_dbl(option1_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_option2 = map2_dbl(option2_mod, test, \(mod, df) rmse(model = mod, data = df)))

# draw the violin plot for rmse
cv_df |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  theme_minimal()

# In cross-validation, the root mean squared error (RMSE) is a measure of the average magnitude of prediction errors made by a model. A lower RMSE indicates that the model has better predictive accuracy, as it means the model's predictions are closer to the true values.

# From the violin plot, we can know that the stepwise regression model has the lowest rmse, while option1 (with two predictors) has the worst rmse. So the stepwise regression model predicts most accurately among three.
```

